Examining the critical path neighborhood for flexible job shop problems

### 2016-01-25 paper/paper.tex
updated paper, run script

\begin{table}[h]
\centering
\small
\begin{tabular}{@{\bfseries}rrrrrrrrrrrrr}
	\toprule
	\multicolumn{4}{c}{Instance} && \multicolumn{2}{c}{Properties} && \multicolumn{5}{c}{Results} \\
	\cmidrule(r){1-4} \cmidrule(lr){6-7} \cmidrule(lr){9-13}
	\normalfont{name} & $|\mathcal{J}|$ & $|\mathcal{T}|$ & $|\mathcal{M}|$ && $\max C_\text{job}$ &
	$C_\text{opt}$ && $\alpha$ & restarts & $\bar{C}$ & $C_\text{best}$ & $\bar{t}$ \\
	\midrule
	ft06 &  6 &  6 &  6 &&   47 &   55 &&   0.99 & 1 &   55 &   55 &    0 \\
	ft10 & 10 & 10 & 10 &&  655 &  930 &&    0.9 & 5 &  954 &  941 &    2 \\
	     &    &    &    &&      &      &&   0.99 & 5 &  943 &  938 &   18 \\
	     &    &    &    &&      &      &&  0.999 & 5 &  933 &  930 &  182 \\
	ft20 & 20 &  5 &  5 &&  387 & 1165 &&    0.9 & 5 & 1215 & 1186 &    1 \\
	     &    &    &    &&      &      &&   0.99 & 5 & 1182 & 1178 &   13 \\
	     &    &    &    &&      &      &&  0.999 & 5 & 1170 & 1165 &  130 \\[2ex]
	c1   &  4 &  2 &  4 &&   34 &   34 &&    0.9 & 1 &   34 &   34 &    0 \\
	c2   &  4 &  2 &  4 &&   53 &   53 &&    0.9 & 1 &   53 &   53 &    0 \\
	c3   &  4 &  2 &  4 &&  133 &  143 &&    0.9 & 1 &  143 &  143 &    0 \\
	c4   &  6 &  2 &  4 &&   52 &   57 &&   0.99 & 5 &   57 &   57 &    2 \\
	c5   &  6 &  2 &  4 &&  105 &  113 &&   0.99 & 5 &  113 &  113 &    2 \\
	c6   &  6 &  2 &  4 &&  205 &  225 &&   0.99 & 5 &  225 &  225 &    2 \\
	c7   &  4 &  3 &  9 &&   39 &   39 &&    0.9 & 1 &   39 &   39 &    0 \\
	c8   &  4 &  3 &  9 &&   64 &   64 &&    0.9 & 1 &   64 &   64 &    0 \\
	c9   &  4 &  3 &  9 &&  138 &  138 &&    0.9 & 1 &  138 &  138 &    0 \\
	c10  &  6 &  3 &  9 &&   50 &   50 &&    0.9 & 1 &   50 &   50 &    0 \\
	c11  &  6 &  3 &  9 &&  103 &  103 &&    0.9 & 1 &  103 &  103 &    0 \\
	c12  &  6 &  3 &  9 &&  204 &  204 &&    0.9 & 1 &  204 &  204 &    0 \\
	c13  &  8 &  3 &  9 &&  337 &  337 &&    0.9 & 5 &  337 &  337 &    0 \\
	c14  &  8 &  3 &  9 &&  705 &  705 &&    0.9 & 5 &  705 &  705 &    0 \\
	c15  &  8 &  3 &  9 && 1458 & 1458 &&    0.9 & 5 & 1458 & 1458 &    0 \\
	c16  & 10 &  4 & 16 &&  405 &  405 &&    0.9 & 5 &  405 &  405 &    0 \\
	c17  & 10 &  4 & 16 &&  955 &  955 &&    0.9 & 5 &  955 &  955 &    0 \\
	c18  & 10 &  4 & 16 && 1621 & 1621 &&    0.9 & 5 & 1621 & 1621 &    0 \\
	c19  & 15 &  5 & 25 && 2631 & 2631 &&    0.9 & 5 & 2631 & 2631 &    0 \\[2ex]
	v20  &  8 &  2 &  4 &&      &      &&    0.9 & 5 &      &      &      \\
	v30  & 16 &  2 &  4 &&  666 &      &&    0.9 & 5 & 2382 & 2381 &   10 \\
	     &    &    &    &&      &      &&   0.99 & 5 & 2381 & 2381 &  102 \\
	v31  & 16 &  2 &  8 &&  566 &      &&    0.9 & 5 &  963 &  961 &   12 \\
	     &    &    &    &&      &      &&   0.99 & 5 &  960 &  959 &  126 \\
	v32  & 16 &  4 &  8 &&  655 &      &&    0.9 & 5 & 1119 & 1116 &   11 \\
	     &    &    &    &&      &      &&   0.99 & 5 & 1112 & 1111 &  116 \\
	\bottomrule
\end{tabular}
\caption{Computational results for the FJS}
\label{table:fjs}
\end{table}

@@

There is little literature on the FBJS, and as a result there are few benchmark instances and results. In any case, these would not be applicable here due to the less generalized formulation we have used. To better assess the critical path neighborhood, we introduce a naive shift neigbhorhood for comparison.

In this neighborhood, we take a random operation of a random machine type, and shift it a random number of (permissable) places left or right. It may also take all of operations of the job this operation belongs to, and shift them all an individually random number of places left. The blocking constraint means that the operations belonging to a job must be executed as close together as possible. In this sense, it makes sense to move these operations as a unit, rather than individually.

@@

Here are some results.

@@

@@

>>>>

\begin{table}[h]
\centering
\small
\begin{tabular}{@{\bfseries}rrrrrrrrrrrr}
	\toprule
	\multicolumn{4}{c}{Instance} & \multicolumn{2}{c}{Properties} && \multicolumn{5}{c}{Results} \\
	\cmidrule(r){1-4} \cmidrule(lr){5-6} \cmidrule(lr){8-12}
	\normalfont{name} & $|\mathcal{J}|$ & $|\mathcal{T}|$ & $|\mathcal{M}|$ & $\max C_\text{job}$ &
	$C_\text{opt}$ && $\alpha$ & restarts & $\bar{C}$ & $C_\text{best}$ & $\bar{t}$ \\
	\midrule
	ft06 &  6 &  6 &  6 &   47 &   55 &&    0.9 & 1 &   55 &   55 &    0 \\
	ft10 & 10 & 10 & 10 &  655 &  930 &&    0.9 & 5 &  954 &  941 &    2 \\
	     &    &    &    &      &      &&   0.99 & 5 &  943 &  938 &   18 \\
	     &    &    &    &      &      &&  0.999 & 5 &  933 &  930 &  182 \\
	ft20 & 20 &  5 &  5 &  387 & 1165 &&    0.9 & 5 & 1215 & 1186 &    1 \\
	     &    &    &    &      &      &&   0.99 & 5 & 1182 & 1178 &   13 \\
	     &    &    &    &      &      &&  0.999 & 5 & 1170 & 1165 &  130 \\[2ex]
	c1   &  4 &  2 &  4 &   34 &   34 &&    0.9 & 1 &   34 &   34 &    0 \\
	c2   &  4 &  2 &  4 &   53 &   53 &&    0.9 & 1 &   53 &   53 &    0 \\
	c3   &  4 &  2 &  4 &  133 &  143 &&    0.9 & 1 &  143 &  143 &    0 \\
	c4   &  6 &  2 &  4 &   52 &   57 &&   0.99 & 5 &   57 &   57 &    2 \\
	c5   &  6 &  2 &  4 &  105 &  113 &&   0.99 & 5 &  113 &  113 &    2 \\
	c6   &  6 &  2 &  4 &  205 &  225 &&   0.99 & 5 &  225 &  225 &    2 \\
	c7   &  4 &  3 &  9 &   39 &   39 &&    0.9 & 1 &   39 &   39 &    0 \\
	c8   &  4 &  3 &  9 &   64 &   64 &&    0.9 & 1 &   64 &   64 &    0 \\
	c9   &  4 &  3 &  9 &  138 &  138 &&    0.9 & 1 &  138 &  138 &    0 \\
	c10  &  6 &  3 &  9 &   50 &   50 &&    0.9 & 1 &   50 &   50 &    0 \\
	c11  &  6 &  3 &  9 &  103 &  103 &&    0.9 & 1 &  103 &  103 &    0 \\
	c12  &  6 &  3 &  9 &  204 &  204 &&    0.9 & 1 &  204 &  204 &    0 \\
	c13  &  8 &  3 &  9 &  337 &  337 &&    0.9 & 5 &  337 &  337 &    0 \\
	c14  &  8 &  3 &  9 &  705 &  705 &&    0.9 & 5 &  705 &  705 &    0 \\
	c15  &  8 &  3 &  9 & 1458 & 1458 &&    0.9 & 5 & 1458 & 1458 &    0 \\
	c16  & 10 &  4 & 16 &  405 &  405 &&    0.9 & 5 &  405 &  405 &    0 \\
	c17  & 10 &  4 & 16 &  955 &  955 &&    0.9 & 5 &  955 &  955 &    0 \\
	c18  & 10 &  4 & 16 & 1621 & 1621 &&    0.9 & 5 & 1621 & 1621 &    0 \\
	c19  & 15 &  5 & 25 & 2631 & 2631 &&    0.9 & 5 & 2631 & 2631 &    0 \\[2ex]
	v20  &  8 &  2 &  4 &      &      &&    0.9 & 5 &      &      &      \\
	v30  & 16 &  2 &  4 &  666 &      &&    0.9 & 5 & 2382 & 2381 &   10 \\
	     &    &    &    &      &      &&   0.99 & 5 & 2381 & 2381 &  102 \\
	v31  & 16 &  2 &  8 &  566 &      &&    0.9 & 5 &  963 &  961 &   12 \\
	     &    &    &    &      &      &&   0.99 & 5 &  960 &  959 &  126 \\
	v32  & 16 &  4 &  8 &  655 &      &&    0.9 & 5 & 1119 & 1116 &   11 \\
	     &    &    &    &      &      &&   0.99 & 5 & 1112 & 1111 &  116 \\
	\bottomrule
\end{tabular}
\caption{Computational results for the FJS}
\label{table:fjs}
\end{table}

@@

There is little literature on the FBJS, and as a result there are even few benchmark instances and results. In any case, these would not be applicable here due to the less generalized formulation we have used. To better assess the critical path neighborhood, we introduce a naive shift neigbhorhood for comparison.

In this neighborhood, we take a random operation of a random machine type, and shift it a random number of (permissible) places left or right. It may also take all of operations of the job this operation belongs to, and shift them all an individually random number of places left. The blocking constraint means that the operations belonging to a job must be executed as close together as possible. In this sense, it makes sense to move these operations as a unit, rather than individually.

@@

The same instances are used as before. A new column, containing the best non-blocking solution for the instance, is added for comparison.

\begin{table}[h]
\centering
\small
\begin{tabular}{@{\bfseries}rrrrrrrrrrrrr}
	\toprule
	& FJS & FBJS & && \multicolumn{2}{c}{Critical Path} &&
	\multicolumn{2}{c}{Naive shift} && \multicolumn{2}{c}{Both} \\
	\cmidrule(lr){2-2} \cmidrule(lr){3-3} \cmidrule(lr){6-7}
	\cmidrule(lr){9-10} \cmidrule(lr){12-13}
	\normalfont{name} & $C_\text{best}$ & $C_\text{opt}$ & $\alpha$ && $\bar{C}$ & $C_\text{best}$ &&
	$\bar{C}$ & $C_\text{best}$ && $\bar{C}$ & $C_\text{best}$ \\
	\midrule
	ft06 &   55 &   63 &  0.99 &&  102 &   95 &&   64 &   63 &&   66 &   63 \\
	ft10 &  930 & 1068 &   0.9 && 2183 & 1931 && 1507 & 1436 && 1464 & 1415 \\
	     &      &      &  0.99 && 2139 & 1931 && 1444 & 1353 && 1413 & 1349 \\
	     &      &      & 0.999 && 2366 & 2028 && 1380 & 1281 && 1371 & 1292 \\
	ft20 & 1165 &      &   0.9 && 2476 & 2406 && 1957 & 1866 && 1935 & 1869 \\
	     &      &      &  0.99 && 2454 & 2348 && 1840 & 1814 && 1860 & 1840 \\
	     &      &      & 0.999 && 5 & 1170 && 1165 &  130 &&   & \\[2ex]
	c1   &   34 &      &  0.99 &&   66 &   62 &&   37 &   37 &&   37 &   37 \\
	c2   &   53 &      &  0.99 &&   88 &   75 &&   65 &   65 &&   65 &   65 \\
	c3   &  143 &      &  0.99 &&  206 &  206 &&  180 &  180 &&  180 &  180 \\
	c4   &   57 &      &  0.99 &&  116 &  109 &&   94 &   92 &&   91 &   90 \\
	c5   &  113 &      &  0.99 &&  247 &  215 &&  180 &  170 &&  175 &  170 \\
	c6   &  225 &      &  0.99 &&  502 &  434 &&  365 &  360 &&  360 &  353 \\
	c7   &   39 &      &  0.99 &&   39 &   39 &&   39 &   39 &&   39 &   39 \\
	c8   &   64 &      &  0.99 &&   64 &   64 &&   64 &   64 &&   64 &   64 \\
	c9   &  138 &      &  0.99 &&  138 &  138 &&  138 &  138 &&  138 &  138 \\
	c10  &   50 &      &  0.99 &&   50 &   50 &&   50 &   50 &&   50 &   50 \\
	c11  &  103 &      &  0.99 &&  170 &  116 &&  106 &  106 &&  106 &  106 \\
	c12  &  204 &      &  0.99 &&  265 &  210 &&  208 &  208 &&  208 &  208 \\
	c13  &  337 &      &  0.99 &&  557 &  524 &&  467 &  458 &&  393 &  389 \\
	c14  &  705 &      &  0.99 && 1203 & 1036 &&  866 &  833 &&  797 &  788 \\
	c15  & 1458 &      &  0.99 && 2556 & 2258 && 1789 & 1666 && 1508 & 1493 \\
	c16  &  405 &      &  0.99 &&  519 &  483 &&  444 &  429 &&  405 &  405 \\
	c17  &  955 &      &  0.99 && 1257 & 1009 &&  962 &  955 &&  955 &  955 \\
	c18  & 1621 &      &  0.99 && 2314 & 2131 && 1710 & 1688 && 1621 & 1621 \\
	c19  & 2631 &      &  0.99 && 4568 & 3973 && 4365 & 4049 && 2631 & 2631 \\[2ex]
	v20  &      &      &   0.9 &&  571 &  540 &&  494 &  476 &&  438 &  423 \\
	     &      &      &  0.99 &&  527 &  517 &&  447 &  421 &&  424 &  418 \\
	v21  &      &      &   0.9 &&  331 &  293 &&  277 &  254 &&  218 &  215 \\
	     &      &      &  0.99 &&  327 &  291 &&  213 &  209 &&  198 &  194 \\
	v22  &      &      &   0.9 &&  419 &  387 &&  326 &  314 &&  294 &  284 \\
	     &      &      &  0.99 &&  393 &  369 &&  293 &  281 &&  273 &  262 \\
	v30  & 2381 &      &   0.9 && 4210 & 4068 && 5969 & 5856 && 3546 & 3438 \\
	     &      &      &  0.99 && 4200 & 4154 && 5591 & 5411 && 3287 & 3147 \\
	v31  &  959 &      &   0.9 && 1658 & 1573 && 4303 & 4133 && 1702 & 1635 \\
	     &      &      &  0.99 && 1688 & 1632 && 3772 & 3712 && 1646 & 1617 \\
	v32  & 1111 &      &   0.9 && 5 & 1119 && 1116 &   11 &&   & \\
	     &      &      &  0.99 && 2629 & 2601 && 1111 &  116 &&   & \\
	\bottomrule
\end{tabular}
\caption{Computational results for the FBJS}
\label{table:fbjs}
\end{table}

@@

It is immediately apparent that for almost all instances, the critical path neighborhood fails to search large portions of the solution space. It quickly converges to a certain makespan, and then fails to improve on it -- even if it is very far off the optimal solution. Sometimes, this makespan is even worse than one reached by the naive shift neighborhood. Neither of the neighborhoods seem to reach good solutions at all, except in the case of an excess of machines. The combination of the two neighborhoods performs better than either of the neighborhoods alone, but the solution quality and speed of convergence still leave much to be desired.

Analyzing the critical path neighborhood reveals a very high percentage of infeasible solutions in the neighborhood. The critical path neighborhood is quite small, which is a strength in the FJS case. However, this means it explores very little of the solution space, and the search quickly becomes stuck due to the large amounts of infeasible solutions. Combining the two neighborhoods improved the search somewhat. In any case, it appears as though moving individual operations on the critical path simply creates infeasible solutions in too many cases.

The blocking constraint means that good solutions will have jobs packed tightly together, in order to minimize the times that machines are blocked instead of processing. Attempting to move one operation out of a tightly packed job to a new time in the schedule will almost inevitably lead to a variable cycle, and an infeasible solution. This is most likely the cause of the failure of the critical path neighborhood when applied to the Blocking Job Shop. It is clear a different approach is needed, perhaps one that looks at entire jobs instead of individual operations.

@@

We have examined the critical path neighborhood in conjunction with a local search for the Flexible Job Shop and Flexible Blocking Job Shop problems.

For the Flexible Job Shop problem, the critical path neighborhood is a very effective search mechanism, performing nearly as well as it does for the classical Job Shop. It doesn't provide the same feasibility guarantees, but in practice this is hardly an issue. With a small number of random restarts, the search will almost invariably find a good solution, and converge very quickly. The strength of this neighborhood lies in its small size and the clever use of the problem characteristics.

These strengths quickly turn into weaknesses when the blocking constraint is added to the problem. The critical path neighborhood fails to effectively search the solution space, and performs more poorly than even the most primitive neighborhoods. Even when combined with a different neighborhood to help it explore different areas of the solution space, it fails to make a significant impact on the solution quality and the rate of convergence. A different approach is needed, one that takes important relationships between operations of the same job into account.

There are several options here. One is to adopt a job removal/insertion technique, similar to one used by Groeflin and Klinkert for the Blocking Job Shop \cite{lsn-bjs}. This is expensive, but ensures feasible solutions. Another is to temporarily allow infeasible solutions. In any case, the representation needs to be extended to account for the blocking constraints, in the form of additional edges. The lack of such a structure in the implementation used in this paper made it exceedingly difficult to recognize and deal with variable or blocking cycles.

To better test new neighborhoods, it's advisable to extend the FJS formulation to the one more commonly found in literature, where operations may be processed on some subset of all machines, instead of non-overlapping machine types. This, combined with benchmark instances, will allow for better comparison and assessment.
